---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-ms.tex
geometry: margin=1in
title: "The Effect of Terrorism on Judicial Confidence"
thanks: "**Corresponding author**: Steven V. Miller, Department of Political Science, Clemson University, 232 Brackett Hall, Clemson, SC 29634-1354. Email: svmille@clemson.edu."
# The author thanks---but cannot blame---Matthew Clary, Chris Kromphardt, Julia West, and the editors and three anonymous reviewers at *Political Research Quarterly* for their comments on previous versions of this manuscript. Replication files are available on the author's Github account ([github.com/svmiller](https://github.com/svmiller)).
author:
- name: Steven V. Miller
  affiliation: Clemson University
  email: svmille@clemson.edu
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: "Independent judiciaries prevent democratic reversals, facilitate peaceful transitions of power, and legitimate democracy among citizens. We believe this judicial independence is important for citizen-level judicial confidence and faith in democratic institutions. I challenge this and argue that citizens living under terror threats lose confidence in their independent judiciaries. Terror threats lead citizens to enable the state leader to provide counter-terrorism for their security, which has important implications for inter-branch relations between the executive and the judiciary. Citizens lose confidence in independent judiciaries that provide due process for suspected terrorists. I test my argument with mixed effects models that incorporate the Global Terrorism Database and four waves of European Values Survey. The analyses demonstrate the negative effects of terror threats on judicial confidence when interacting terror threats with measures of judicial independence. My findings have important implications for the study of democratic confidence and the liberty-security dilemma."
fontfamily: mathpazo
fontfamilyoptions: sc, osf
fontsize: 11pt
keywords: "terrorism, judicial confidence, judicial independence, liberty-security dilemma"
bibliography: ~/Dropbox/master.bib
biblio-style: apsr
anonymous: no
endnotes: yes
---

```{r loadpackages, echo=F, message=F, warning=F, error=F}
setwd("~/Dropbox/projects/etjc")

library(car)
library(arm)
library(tidyverse)
library(countrycode)
library(stringr)
library(mirt)
library(stargazer)
library(AICcmodavg)



theme_steve <- function() {
       theme_bw() +
  theme(panel.border = element_blank(), 
        plot.caption=element_text(hjust=1, size=9,
                                               margin=margin(t=10),
                                               face="italic"),
        plot.title=element_text(hjust=0, size=18,
                                             margin=margin(b=10),
                                             face="bold"),
        axis.title.y=element_text(size=12,hjust=1,
                                                face="italic"),
        axis.title.x=element_text(hjust=1, size=12, face="italic"))
  
}

```

```{r loaddata, echo=F, message=F, warning=F, error=F, cache=T}
source('analysis/1-load-data.R') 
```

```{r cleanmicro, echo=F, message=F, warning=F, error=F, cache=T}
source('analysis/2a-clean-evs-micro.R')
```


```{r cleanmacro, echo=F, message=F, warning=F, error=F, cache=T}
source('analysis/2b-clean-evs-macro.R') 
```

```{r analysis, echo=F, message=F, warning=F, error=F, cache=T}
source('analysis/3-analysis.R')
```



# Introduction


Why is confidence in the judicial system so low in a country like Great Britain? British popular confidence in the judicial system lags behind both British confidence in other state institutions [@vandewalle2009ccjs] and lags well behind most of its democratic peers in Western Europe [@buhlmannkunz2011cj]. A comparison of European survey data places British confidence in the judicial system closer to the mass-level confidence in judicial systems in non-democracies like Azerbaijan and Ukraine than democratic peers like Denmark or Iceland. This is curious because the British judicial system shows numerous benchmarks that should inspire widespread public confidence. The ability of the judicial process to resolve commercial disputes and enforce contracts places the British system as No. 6 in the world for judicial quality and efficiency [@wbg2016db]. Its time to resolution from submission of a claim is among the best in Europe and the world. Civil liberty and due process protections are also strong in the United Kingdom. Its protection of political rights and civil liberties gives the United Kingdom a 95 of 100 in the 2016 *Freedom House* rankings [@fh2016fw], which ties United Kingdom with many of its democratic peers for No. 6 in the world. Judicial independence data [e.g. @cingranelli2014ciri<!--; @riosfigueroastaton2014ecnm-->] give further confidence that the British judicial system is independent from the influence of other state or societal actors. Yet, confidence in the British judicial system lags behind confidence in other British institutions and judicial confidence elsewhere in Western Europe.

The independence the British judicial system enjoys may actually be part of the problem. The United Kingdom, unlike Denmark and Iceland, has a long history of terror threats. This had historically been the Irish Republican Army though recent threats involve Islamist terror groups like Al-Qaeda or the Islamic State. The British judicial response to these terror threats stands at odds with executive-led initiatives in countries like Australia and the United States after September 11, 2001 [@roach2011cct] or the symbiotic executive-judiciary relationship that France had implemented in 1986. The British judicial system institutes numerous legal protections for terror suspects---including automatic legal representation during interrogations and a high threshold for charging a suspect with terror-related offenses---that diverges from the French model of counter-terrorism. British politicians even express envy at the more "flexible", security-oriented judicial response to terrorism in France [@foley2013ctbf]. It is not uncommon for British politicians to criticize their independent judges, nor for tabloids in the United Kingdom to blast judges on the front page as "out of touch" for the legal protections they provide to terror suspects [e.g. @dyer2005jso; @moore2015otj]. This suggests an answer to the puzzle of sagging confidence in the British judicial system. The British may be losing faith in a judicial system as independent and civil liberty-oriented as the British model because terror threats lead citizens to prefer administrative or executive solutions to deal with terror threats. Citizens perceive independent judiciaries as interferences to national security.

In this paper, I generalize from the British case and argue a general relationship between terror threats and individual-level confidence in the judicial system across four waves of European Values Survey data. I argue that terror groups engage in attacks that target citizens and use violent fear tactics to coerce support for policies in line with the terror group's preferences. Citizens targeted by these terror attacks offset the subsequent fear these attacks instill by rallying around their leaders to provide for their security. This support for state leadership provides leverage in inter-branch relations as rally effects induce acquiescence from political rivals to the leader's agenda. I show this extends to individual-level evaluations of the judiciary as well. Under conditions of high terror threat, citizens lose confidence in their judicial system that they interpret as providing due process and legal protections for suspected terrorists and terror groups. Paradoxically, we observe the negative effects of terror threats on judicial confidence the *more* independent the judicial system is from other branches of government. This interactive effect allows for governments to reintegrate more judicial processes into the executive branch like we are currently observing in France after *Charlie Hebdo*, and have observed in the United States since 2001.

The rest of this paper starts with a review of the importance of independent judiciaries to democratic governance and what explains confidence citizens have in these important institutions of government. Thereafter, I outline my theory about how terror threats negatively affect individual-level confidence in the judicial system and why this effect is contingent on the independence of the judiciary from other branches of government. This leads to the next two sections that detail my research design and the results of mixed effects models of four waves of European Values Survey data and 29 years of Global Terrorism Database data. I conclude with implications of my analyses.


# Judicial Independence and Confidence: The Importance of Two Concepts


Judicial independence is a multidimensional concept and has multiple meanings in different contexts [@russellobrien2001jiad]. @shapiro1981c sees judicial independence as institutional, defining it as the existence of a neutral third party to serve as impartial adjudicator over conflicts among societal actors. @larkins1996jid extends this idea of impartiality but suggests judicial independence concerns norms as well. In his view, judicial independence exists when there is an institution that regulates the legality of government behavior but can also determine constitutional or legal values. Recent conceptual definitions reinforce judicial independence as multidimensional with "de jure" and "de facto" subconcepts [e.g.: @feldvoigt2003egji<!--; @riosfigueroastaton2014ecnm-->]. "De jure" judicial independence refers to the extent to which formal rules exist to insulate judges from political pressure. In this context, judicial independence exists when judges have benefits like fixed tenure or a level of budgetary autonomy that serve as institutional barriers between their position and other political actors. "De facto" judicial independence is a more behavioral component to the overall concept. This approach suggests an independent judiciary exists when judges adjudicate on the law in a way that reflects their sincere opinions and that other institutions in society enforce their decisions.

Three themes recur when we determine whether a judiciary is "independent" even as scholars offer different interpretations of what judicial independence entails [e.g. @fiss199394lji<!--; @stephenson2009ji-->]. One, an independent judiciary should be impartial, whether "impartiality" means that judges do not have personal interests in cases they adjudicate or that judges are not appointed to their positions because of their political views. Two, an independent judiciary is an institution whose decisions on conflicts among societal actors are implemented and respected even if implementation requires coercion from some other societal actor. Finally, an independent judiciary is insulated from pressure by other societal actors who have personal interests in an issue under adjudication. 

Social scientists and legal scholars invest energy in all that entails "judicial independence" because there is a near consensus that an independent judiciary is essential to prosperity and democratic survival. Consider first the economic benefits of an independent and coherent judiciary body. @landesposner1975ijig conceptualize judicial independence as institutions that maximize the value of legislative bargaining among interest groups by enhancing the durability of the deals to which they agree. In other words, independent courts and judicial officials are credible commitment devices that provide assurances that a contract signed today will endure tomorrow. The benefits that follow are multiple. Independent judiciaries protect investor rights and economic freedoms that induce more economic activity and investment [@laportaetal2004jcb]. This leads to national economic growth [@feldvoigt2003egji]. Evidence from post-Glorious Revolution England [@northweingast1989cc] and post-Soviet Russia [@frye2004ccpr] underscore the need for strong independent courts that protect property rights and promote investment necessary for growth.

Judicial independence may be more important for democratic survival even if opinion differs on what role independent courts play in democracy [@giblerrandazzo2011teij].<!--[^dissents]--> Judicial annulment---i.e. the declaration of an executive decision or legislation as impermissible under the current constitutional charter---is the most visible role that independent judiciaries play in democracies, though this is just one example. @stonesweet2000gj's analysis of parliamentary politics in Europe highlights a consultative role the judiciary plays in the legislative process. Rather than risk judicial annulment, legislators consult with judges about what terms of a bill are constitutionally permissible and revise legislation accordingly. Thus, an independent judiciary influences the legislative process in a manner that protects democratic principles outlined in constitutional charters. Others make even stronger arguments about the importance of judicial independence to democracy. @larkins1996jid argues judicial independence protects minority rights consistent with an interpretation of a liberal constitution. An even stronger argument identifies the rule of law provided by independent and impartial judicial systems as the link that connects democracy and peace [@oconnor2008fic]. This conforms well to @linzstepan1996tcd's contention that independent judiciaries are necessary for democracies to consolidate and endure after transitions of power from one regime to another.

<!-- [^dissents]: Do see @dahl1957dmd and @tsebelis2002vp for dissents of this perspective. -->

Judicial independence carries important implications for citizen confidence in democracy. For one, the importance of judicial independence to economic prosperity concerns scholarship on attitudes toward democracy. We know from @inglehartwelzel2005mccd that there is almost no distinction between citizens who support democracy and citizens who do not, only a distinction between those who support democracy for its own intrinsic value and those who support democracy the extent to which it provides order and prosperity. This finding about attitudes in favor of democracy stands at odds with the actual distribution of democracy across the globe. The implication is institutions that fairly adjudicate disputes and enforce contracts allow for economic activity and growth, fostering confidence in democratic governments. Whereas independent judiciaries protect minority rights [c.f. @larkins1996jid], this suggests that this protection enhances popular legitimacy of democracy among those minority groups. Thus, these groups are less likely to rebel against majoritarian, if democratically elected, governments. Ultimately, literature concerned with popular confidence in democracy requires understanding popular confidence in the judiciary as a critical element of democracy. Even if democracy can still thrive as a function of *mistrust* and a lack of confidence among citizens in the executive and legislature [@norris2011dd], a lack of confidence in the judicial system casts doubt on democracy's trajectory in a given country.

What explains when citizens have confidence in their judicial system given its importance to democracy? Intuitively, citizens have confidence in their judicial system when they know that their judicial system is independent [@buhlmannkunz2011cj]. A judicial system that is independent from the influence of other societal actors fosters confidence in the typical citizen that the judicial system is adequately providing rule of law and fair adjudication. Further, uncertainty in the judicial process decreases confidence in the justice system [@dakoliassaid1999jr]. Judicial systems in which adjudication processes are protracted or are subject to frequent delays decrease individual-level confidence in the judicial system as a whole. This motivates part of the World Bank's *Doing Business* project, which assesses, in part, how resolute and streamlined judiciaries are in the enforcement of contracts [see: @djankovetal2003c]. The conclusions here are rather straightforward. Citizens lose confidence in the judicial system when it is not doing the job expected of it in a fair, impartial, speedy, and thorough manner.

However, such a simple conclusion about the sources of judicial confidence belies policy changes we have observed in some democracies in the War on Terror era. Citizens in these countries appear to be losing confidence in their judicial system despite what otherwise looks to be an efficient and independent judiciary in a comparative perspective. The particulars of American public opinion on this topic are surprising. @huddyetal2005tas find that citizens surveyed from October 2001 to March 2002 who had a high perceived threat of a future terrorist attack were much more likely to support Bush-era policies at the heart of the War on Terror. The effects were not just directed at Arab-Americans as an in-group/out-group effect. Importantly, they were more likely to support government monitoring of telephones and e-mails and were more likely to rescind their own civil liberty benefits for the cause of security. @hetheringtonsuhay2011ata find a similar effect for perceived terror threats on warrantless wiretapping, torture, media censorship, and broad public surveillance measures. All the policies that citizens support run counter to the legal protections provided by independent judiciaries in democratic systems. In other words, American citizens worried about terror threats are likely to support policies that infringe upon civil liberties that independent judicial systems are supposed to protect. Whereas the United States consistently ranks near the top among democracies for judicial efficiency measures gathered in the *Doing Business* project by the World Bank, public support for these policies indicates decreasing judicial confidence as a function of terror threats.

In the next section, I outline a theory that links terror threats to decreasing judicial confidence among citizens.


# Terror Threats, Terror Management, and Attitudes toward the Judicial System

My theoretical argument starts with a definition of terrorism and what these acts try to accomplish. Per @enderssandler2012pet [4], terrorism is the premeditated use or threat of violence by individuals or subnational groups to obtain some objective that is political or social in nature. The threat or use of violence is intended to intimidate a larger audience beyond the immediate victims. Two themes recur that are captured in this definition. The threat or use of violence is necessary to get a political decision-maker to respond to the demands of the individual or group that initiate the acts. It is also necessary that the motive for the terrorists be political or social in nature. Without a political or social motive, violent acts are conceptually indistinguishable from crime and typically do not have the pervasive effects beyond the immediate victims.

Terrorism scholars split hairs about issues related to the identity of the perpetrator and the immediate victim of a suspected terror attack. However, scholars almost universally have the same conceptualization of the identity of the "larger audience." Terrorists design these acts to intimidate the public at large and instill in them a heightened level of anxiety and fear. They do this for two reasons. One, terrorists may have found it difficult to achieve their desired political or social objectives within the normal (i.e. state-sanctioned) political procedures. Therefore, they pursue their objectives with the threats or the use of violence outside normal channels. Two, terrorists want to intimidate the general public as part of a three-player bargaining game over a current public policy for which terrorists seek revisions. Terrorists use these acts to signal to the general public that the costs associated with public support for a government policy the terrorists want to revise are greater than the general public had anticipated. In this interpretation, a successful terror campaign<!--, like the canonical case of the 1983 Beirut barracks boming [@pape2003slst],--> frightens the general public, encouraging them to apply pressure to political decision-makers to concede to the terrorists' demands. For terrorists, targeting the general public as the "larger audience" and instilling in them a heightened level of anxiety and fear is an important design feature of terrorist acts. <!-- The terrorists' goal is to translate this anxiety and fear into pressure on policy-makers for concessions to the terrorists in exchange for greater security and peace of mind for the general public. -->

Terror management theory<!-- [e.g. @greenbergetal1990etmt]--> gives us a roadmap to how the fear induced by these terror attacks translates to changes in the typical citizen's political attitudes and policy preferences. Terror attacks heighten the salience of mortality as the terrorists intend, though this does not lead to demands among citizens to acquiesce to the terror group's demands. Instead, terror management scholars demonstrate that citizens first suppress concerns of their own mortality and then mount a vigorous defense of their worldview when their mortality is made salient. Terror management theory is versatile and capable of explaining other forms of "terror" but this theory has already demonstrated clear applications to political attitudes in the War on Terror era [e.g. @pyszczynskietal2003iw911].

The vigorous defense that citizens mount for their worldview changes attitudes toward executive authority [@landauetal2004due]. Citizens work to reinforce their worldview through support of leadership---state leadership in this context---after they suppress the fear of their own mortality that terror attacks induce. State leaders provide two essential services for citizens after a terror attack. First, state leaders comfort citizens when they offer a symbolic protection against the source of terror threats. This symbolic protection involves the use of messianic or provocative rhetoric like George W. Bush's proclamation to "rid the world of evil-doers" <!--[@perezrivas2001bvr]--> shortly after September 11, 2001. Second, state leaders provide actual, physical safety against terror threats in the form of counter-terrorism strategies. The counter-terrorism strategies that leaders provide and citizens want are designed to be preemptive, which differs from an ex post punitive strategy that ordinary law enforcement provides [@posner2006nsp]. This maps well to citizens demand for safety from terrorism. True security from the destructive costs of terrorism requires the prevention of their occurrence, not means to hand out punishments after a terror attack has taken its toll. Terror threats lead to individual-level support for executive authority to provide symbolic and physical security from the threat to their mortality.

This amounts to a multifaceted "rally effect," which is more than a simple upturn in approval ratings for the head of state. Citizens who turn to executive authority and empower the position to provide for their security see their leaders less as polarizing political figures and more as personifications of the national ethos (i.e. "the flag") [@lee1977raf]. This also has clear implications for citizen attitudes toward the leader's legislative rivals. Citizens become intolerant of dissidence against the head of state who physically and symbolically constitutes a defense of their worldview. We have already seen the policy implications of this multifaceted rally effect as a result of terror threats, especially in the form of legislative empowerment of executive discretionary powers in democracies across the globe [@epifanio2011lrit]. Legislative rivals are unlikely to challenge the head of state's counter-terrorism agenda and are more likely to put aside partisan political differences and work toward common security with the head of state when there is a common threat like terrorism. The case of post-9/11 United States is again illustrative. Russ Feingold's (D-WI) attempts to amend and delay the USA PATRIOT Act found little support among even his Democrat colleagues. Feingold was the lone "nay" vote to a sweeping counterterrorism legislation pushed by a Republican White House and approved just 36 days after the bill's introduction to a Democrat-controlled Senate. 

Our focus on the implications of terror-induced rally effects for executive-legislative relations or partisan differences in government has missed the important implications of these terror-induced rally effects for inter-branch relations between the executive and judiciary. The executive begins to co-opt basic judicial processes for the cause of counter-terrorism and security as we have seen during the War on Terror [@roach2011cct]. Post-9/11 United States is again illustrative, especially the infamous military prison at Guantánamo Bay in Cuba. This executive-led policy amounts to indefinite detention of suspected terrorists in a system that does not follow domestic criminal laws nor conforms to international laws regarding war. Other democratic countries have pursued executive routes to counter-terrorism that circumvent traditional laws, blur boundaries between crime and war, and bestow executive authority with judicial processes. Canada and Australia, for example, relied on executive-level counter-terrorism measures in addition to administrative detention. Canada strengthened an immigration law in November 2001 to allow for indeterminate detention of non-Canadians perceived as security threats. Australia enacted a series of legislations that bestowed the executive with powers of preventative arrests, control orders, and regulation of perceived terrorist speech both as a result of 9/11 and terrorist attacks directed against Australian nationals as a result of the peacekeeping operation in East Timor. 

The extent to which executive-led counter-terror policies initiated under high terror threats infringe upon judicial processes guaranteed by law and the constitutional charter creates an issue for an independent judiciary in how to respond. This is the familiar problem of "judicial deference" in the public law literature, well-known to those working on the topic of administrative rule-making [e.g. @schiller2007ed<!--; @eskridgebaer2008cd-->] with more recent applications to national security policy during the War on Terror [e.g. @chesney2009nsfd<!--; @pearlstein2010ad-->]. Numerous arguments, both within the judicial system [e.g. @posner2006nsp] and outside it [e.g. @posnervermeule2007tb], contend that it is normatively desirable for courts to defer to the executive under these conditions even if the executive policies that follow compromise civil liberties guaranteed by a liberal constitutional charter. Judges or legal scholars that make these normative arguments typically support their claims by reference to the expertise and endowment advantages of executive authority relative to the courts. Executive authority has expertise on a salient policy concern like national security where judges lack such competence and insight. Thus, judicial deference is desirable because judicial interference stands at odds with why executive authority is empowered and expanded in the first place, i.e. to craft policies that address with the complexity of modern-day societal and political problems [c.f. @landis1938ap]. These arguments also highlight additional endowment advantages of flexibility, power, and secrecy for executive authority. Executive authority is better positioned to quickly and forcefully address these threats, using private information about the nature of its capabilities vis-à-vis the adversary to provide effective counter-terrorism. Courts are slow and rigid by contrast and judicial transparency even compromises an effective counter-terrorism policy that may depend on the ability of executive authority to operate in secrecy or conceal sensitive information. A powerful and independent judiciary that threatens judicial review against perceived executive oversteps does more harm than good in this argument, which punctuate arguments inside and outside the legal system for judicial deference to the executive under conditions of high terror threats  [c.f. @bryankromphardt2016pops].

Incentives for deference in the judicial system are strong in these conditions. Courts are inclined to accept the constitutional validity of executive-led security policies and treat disputes between the executive and other political actors as "political questions" rather than constitutional questions [@henkin1997faus]. High terror threats also compound the judiciary's difficulty in defining and limiting the scope of executive authority from constitutional charters, many of them centuries-old, that are routinely more specific about legislative powers than they are about enumerated executive powers [@craigtomkins2006epl]. However, judicial deference is never complete as the cases reviewed in @craigtomkins2006epl show. The judiciary tends to defer on the broader counter-terrorism strategy and the overall legislation that governments craft as policy responses. The judiciary is more open to challenges to particular aspects of the legislation or conspicuous applications of it. In the British case, this was the legal challenge (*A and Others v. Secretary of State for the Home Department* [2004]) to the power of indefinite detention without trial as contained in one part of the Anti-terrorism, Crime and Security Act passed in the aftermath of the September 2001 terrorist attacks in the United States. The result of that case ruled this part of the 2001 law was incompatible with the European Convention on Human Rights. The U.S. likewise saw numerous legal challenges to particular applications of powers given to the Bush White House. The court decisions that followed in *Hamdi v. Rumsfeld* (2004), *Rasul v. Bush* (2004), *Hamdan v. Rumsfeld* (2006), and *Boumediene v. Bush* (2008) reaffirmed habeas corpus rights for detainees and access to federal courts to appeal detentions. Notice the applications of judicial oversight are narrow and case-specific even in conditions of judicial deference. The *A* case did not invalidate the entirety of the 2001 anti-terrorism law in the United Kingdom nor did the *Hamdi* decision rule that the plaintiff's detention was illegal. An independent judiciary exercising broader judicial deference to counter-terror laws will still exercise judicial oversight on particular topics or applications.

Even light judicial oversight into executive counter-terror policy becomes unpopular, as the decisions in the British and American cases highlight. Tabloids blasted British judges as "out of touch" after the *A* decision [e.g. @dyer2005jso]. In the U.S. case, Supreme Court approval ratings plunged in the wake of the *Hamdi* decision [c.f. @gallup2016idsc] while approval ratings for the Bush White House were unaffected [@wheeler2008ppa, 47-8]. Consistent with public opinion scholarship at the time, terror threats led U.S. citizens to support executive-led policies that infringe on judicial process even for U.S. citizens [e.g. @huddyetal2005tas; @hetheringtonsuhay2011ata]. Applications of judicial oversight, however light within the bigger framework of judicial deference, lead to public frustrations with an independent judiciary and decreasing confidence in the institution that was otherwise providing important protections for civil liberties and due process guaranteed by the law and the constitutional charter.


Counter-terrorism policies in stable democracies that constitute executive encroachment on judicial processes and the public reaction to even light judicial oversight into the application of these policies suggest a decrease in confidence in judicial process among citizens and policymakers. Citizens interpret judicial oversight into counter-terror policies as providing terror suspects with unwanted legal protection and due process that they feel will compromise security. Policymakers advocating for these policies will even make this connection explicit. Philip Ruddock, Australia's attorney general from 2003 to 2007, argued that the broad counter-terrorism measures the Australian government was pursuing were necessary because the best a judicial option could offer was murder convictions after a successful terrorist attack. This is a strong claim to make, but several senior justice officials in Australia supported this contention and the delegation of more judicial power to the executive branch for the cause of national security [@lynchetal2007llwt]. 

Notice the interaction between judicial independence and terror threats in this discussion. The largest reactions against judicial approaches to counter-terrorism have occurred in countries with durable, strong, and independent judiciaries. The policy changes and attitudes of citizens and politicians after September 11, 2001 are different in advanced democracies than other forms of government in which the judiciary is not divorced from other government branches. @roach2011cct notes the contrast between the Australian and U.S. cases (for example) and observations from some Middle East countries where terror threats are more prominent. Egypt and Syria, before their recent revolutions, had already co-opted judicial process for terror threats into executive measures like military courts. Syria's military courts attract attention from human rights groups because these military courts featured a military judge to administer proceedings, complemented by two civilian judges, with no appeal process except to the head of state<!-- [@hrw2009fj]-->. Thus, these two states had no major policy reactions to the 2001 terrorist attack against the World Trade Center. Yet, judicial confidence remained high. 87.3% of Egyptian respondents in the second wave of Arab Barometer had medium or great confidence in their judiciary on a response coded identical to the European Values Survey item I use in my analysis [@ab2011ii]. This would rank Egypt as having the highest level of citizen confidence in the judiciary among countries in the fourth wave of my analyses (see Table A.1 in the appendix). It would rank just ahead of terror-free Denmark and, unsurprising for my hypothesis, a Turkish state under considerable terrorist threat.

Even Israel has seen many judicial processes co-opted from criminal courts and bestowed in military courts to the detriment of judicial independence. State leaders are unencumbered by an independent judiciary in the provision of counter-terrorism and security under these conditions. Israel's case is peculiar because  it has been a competitive parliamentary democracy through its history though the CIRI Human Rights Dataset [@cingranelli2014ciri] codes Israel as having only a "partially independent" judiciary through various stretches of its recent history, including the onset of the Second Intifada in 2000. A 2000 survey revealed 74% of Israelis had high confidence in their judiciary [@mautner2011lci, 164] notwithstanding its lack of judicial independence. This would rank Israel as having the second highest level of citizen confidence in the judiciary among countries in the third wave of my analyses (see Table A.1 in the appendix), behind only terror-free Denmark.


This leads to a testable hypothesis I evaluate in the remainder of the manuscript. The effect of terror threats on judicial confidence in a given year depends on the value of judicial independence. Citizens in states subject to high levels of terror threats are likely to have confidence in their courts provided the judicial system is not independent from other political actors. States in these conditions have already co-opted judicial processes in the form of military tribunals, which leads to citizen confidence that the legal system is working with the executive to provide for the citizen's security. Citizens in states with high levels of terror threats *and* independent judiciaries do not have faith in their judicial system by contrast. Citizens interpret these indpendent judicaries as providing basic legal protections for terror suspects that they do not want and view as an impediment to their own security.

\begin{hypothesis}
\begin{minipage}[t]{5.1 in}
Citizens living in states with high levels of terror threats and judicial independence are less likely to have confidence in their judicial system than citizens in states with high levels of terror threats and without independent judicial systems.
\end{minipage}
\end{hypothesis}

The next section proceeds with a research design to test this hypothesis.

# Research Design

My sampling frame consists of all European countries that appear in the European Values Survey in four waves that span from 1982 to 2009. A supplemental appendix contains additional information about summary statistics, the countries and regions that appear in the analyses, and robustness checks.

## Dependent Variable

I proxy the concept of judicial confidence with a question that has appeared in all waves of European Values Survey (EVS) data as part of its battery of questions about confidence in domestic institutions. This survey item prompts citizens to say how much confidence they have in their judicial system on a four-point ordinal scale of "a great deal", "quite a lot", "not very much", and "none at all". I condense this ordinal measure to a binary indicator and assign a one to those respondents who have "a great deal" or "quite a lot" of confidence in their judicial system.[^ordinal]

[^ordinal]: I also estimate an ordinal logistic mixed effects model of what I present as Table \ref{tab:regtable} in the manuscript. This is available in the appendix.

## Macro-level Independent Variables

My primary interests lay in the effect of contextual (i.e. macro-level) influences on individual-level judicial confidence, especially a relationship between country-level terror threats and judicial confidence that is contingent on the level of judicial independence. I start with a discussion of these two macro-level variables before I describe the macro-level control variables.

I draw my country-level measure of terror threats from the August 2014 update to the *Global Terrorism Database* (GTD) [-@gtd2014] database, which codes all terrorist incidents between 1970 and 2013<!-- (START, \citeyear{gtd2014}) -->. While this data set is a great resource for specific terror incidents since 1970, it may not capture the concept of "terror threats" that may linger for months and years after a terror attack. Further, the third and fourth waves of EVS data have considerable gaps between measurements, for which a general index capturing long-term terror threats would be useful. I thus mirror the @gti2014 <!-- Global Terrorism Index' (\citeyear{gti2014}) --> methodology for coding terror levels, which itself draws data from the Global Terrorism Database (GTD). I first subset the GTD to just countries in Europe rather than the whole international system. I add the number of terrorist incidents in a country in a given year (multiplied by one), the number of civilians killed in those incidents (multiplied by three), the number of civilians injured but not killed (multiplied by .5), and the property damage on GTD's ordinal scale (multiplied by two). This creates a raw terror threat score for a European country in a given year. I then create a five-year weighted average of this raw score to account for lingering effects of terrorism. The current year's raw terror threat score is multiplied by 16, which decrements by half until the raw terror threat score from four years ago (which is multiplied by one). This generates a terror threat index that I log to control for skew, lag to circumvent endogeneity concerns, and group-center (on the survey wave) and scale by two standard deviations [see: @gelman2008sri].[^myriad]

[^myriad]: The appendix contains ten different estimations of the concept of "terror threats," all of which produce identical results for the parameters for judicial independence, terror threats, and the statistical interaction between them. The results I report here do not depend on how I code the terror threat variable.

I code country-level judicial independence with the April 2014 release of the CIRI Human Rights Dataset [@cingranelli2014ciri]. Scholars know this data set well for its use in human rights scholarship but it has considerable use for scholarship related to judicial independence. This data set has a three-part ordinal measure of judicial independence for a country in a given year. CIRI codes a country's judiciary as generally independent if it meets multiple conditions. These include the right to rule on the constitutionality of executive and legislative policies, tenure protections, and the power of judicial annulment. Country judiciaries are partially independent or not independent if these powers are subject to active interference from other government branches or corruption and intimidation from non-governmental actors. I maintain the ordinal nature of this variable where a value of 0 indicates judicial dependence, 1 indicates partial judicial independence, and 2 indicates general judicial independence.

I also consider several control variables at the macro-level. I add dummies for Western Europe and post-9/11 observations as checks for regional differences in Europe and differences between observations in the War on Terror era and those before it. I control for the overall level of democracy in a country with the mean democracy estimate for the country-year in the Unified Democracy Scores (UDS) [@pemsteinetal2010dc] data. @buhlmannkunz2011cj find some evidence that judicial confidence among citizens is concentrated higher in richer countries. Thus, I code the real gross domestic product (GDP) per capita of the country prior to the EVS survey year [@gleditsch2002etgd]. Finally, I control for the possible confounding effect of judicial inefficiency on the relationships among judicial confidence, judicial independence, and terror threats. I use the World Bank's *Doing Business* project to code the average duration (in days) from the time a plaintiff decides to file the lawsuit in court until the final restitution of the claim [see: @djankovetal2003c]. Longer durations signal less efficient and capable judicial systems that decrease judicial confidence and may confound the partial and interactive effects of judicial independence. Though this data source is unique, it has observations starting only in 2004. Thus, use of this important control variable restricts analyses to just the fourth wave of EVS data in one of the models shown in Table \ref{tab:regtable}. All variables at the macro-level were group-centered on the survey wave and scaled by two standard deviations to provide a rough comparability of coefficients in the regression table I present [see: @gelman2008sri].

## Micro-level Control Variables

Though my hypotheses concern macro-level effects on individual-level attitudes of judicial confidence, I specify several micro-level control variables from EVS as well. My collection of individual-level controls mirrors how @buhlmannkunz2011cj conduct their analysis on the same survey item I estimate as my dependent variable. I gather data on the respondent's age, gender, education level, income group, employment status, and confidence in parliament. Age is measured in years. Gender is a binary indicator for those who self-identify as women. Education-level is a binary indicator for those holding a college degree. I measure income with the EVS' subjective income group variable that assigns respondents into "low", "medium", or "high" income brackets contingent on responses to other economic status items that vary across survey waves. Employment status is a binary indicator for those who unemployed. The confidence in parliament question is patterned in a similar fashion to the response variable I analyze though I maintain the four-part ordinal nature of the variable. All variables that are not binary are centered on the country-wave and scaled by two standard deviations to provide a rough comparability of coefficients shown in Table \ref{tab:regtable} in the next section.

# Results

Table \ref{tab:regtable} shows the results of the mixed effects models that estimate individual-level attitudes of judicial confidence. I start with a discussion of Model 1.

```{r regtable, echo=F, results="asis"}
cat(Regtable, sep='\n')
```



Model 1 estimates a minimal macro-level model without an interaction between terror threat levels and judicial independence. The results show no independent partial effect for either judicial independence or terror threat levels on individual-level judicial confidence. The null findings in Model 1 are prima facie consistent with the hypothesis that the effect of terror threats depends on the level of judicial independence in the country. This minimal macro-level model results in only one regression parameter discernible from zero. The Western Europe dummy is positive and significant in Model 1 and even the remaining models, suggesting important heterogeneity between Western Europe and the rest of the EVS countries over four waves of EVS data. This is the only regression paramater significant in Model 1 at the .05 level.


Model 2 adds an interactive term between terror threat levels and judicial independence. The results here are consistent with the hypothesis' prediction of a relationship between terror threat and judicial confidence that is contingent on the independence of the judiciary. The coefficient for the interaction is negative and statistically significant. Citizens in states with higher levels of terrorism and in countries whose judiciaries are independent from other political or societal actors have less confidence in the judicial system and judicial process. Consider the interpretation of the coefficient for terror threat levels as well. The coefficient for terror threat levels in the interaction model specifies the estimated effect of increasing terror threats on citizen attitudes toward the judiciary when the judicial independence variable is zero (i.e. judicial dependence on other political institutions). Therefore, we would conclude that citizens in countries with higher terror threat levels and without independent judiciaries generally have more confidence in their judicial system. 

Model 3 adds the individual-level control variables similar to those specified by @buhlmannkunz2011cj. Only the coefficients for age, female, and confidence in parliament are discernible from zero. Older citizens and less likely to have confidence in their judicial system though women are more likely to have confidence in their judicial system. Citizens who have more confidence in their legislative branch of government are also likely to have confidence in their judicial system as well. Whereas I scale all variables by two standard deviations to approximate a same scale among binary and non-binary predictors, we can have some confidence at first glance that the effect of confidence in parliament is the largest of all the predictors, and certainly the most precise. Importantly, the inclusion of micro-level control variables do not meaningfully influence the coefficients for the macro-level variables. The interaction between terror threats and judicial independence is still negative and statistically significant. This is consistent with my theory that citizens lose faith in their judicial system and judicial process because citizens perceive that judicial process provides protections for suspected terrorists.

Model 4 adds an important macro-level control variable for judicial efficiency. While judicial efficiency may confound the relationship between the independent and dependent variables of interest, our best available data for judicial efficiency (from the World Bank’s *Doing Business* project) come only after 2004. This restricts the analyses to just the fourth wave of EVS data and drops the post-9/11 dummy from the model. Model 4 shows a significant effect of judicial inefficiency—as measured in average calendar days from a plaintiff’s lawsuit until the final restitution of the claim—and judicial confidence. Longer durations suggest less efficient judicial systems, which leads to less confidence among citizens in the judicial process. However, the inclusion of judicial inefficiency as an important correlate and the restriction of the analyses to just the fourth EVS wave do not meaningfully affect the parameter estimates associated with judicial independence, terror threat levels, and the interaction between the two. Citizens in states with high levels of terrorism and independent judiciaries are less likely to have confidence in their judicial system.


```{r graph, echo=F, results="asis", fig.cap="The Interaction between Judicial Independence and Terror Threats on Judicial Confidence", fig.width=8, warning=F}
ggplot(M2_predvalues, aes(x = zg_loggti5ya, y = pp)) + 
  theme_steve() +
  geom_errorbar(aes(ymin = lb, ymax = ub), alpha = 0.3, linetype = 1) + 	
  geom_line(aes(colour = Category, linetype = Category), size = 1) +
  ylab("Predicted Probability of Judicial Confidence") +
  xlab("Range of Terror Threat") +
  theme(legend.position="bottom") +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by=.1)) +
  scale_x_continuous(breaks=c(seq(-1.1,1.1, by=.1)), limits=c(-1.1,1.1))

```


Figure 1 displays quantities of interest from Model 2 and presented in Table \ref{tab:regtable}. I set all but the variables of interest at their typical value, for each independent variable shown in Model 2. Then, I generated 300 observations across the entire range of the terror threat variable, repeated twice. I code half of the observations on the judicial independence variable with the minimum of zero (i.e. no judicial independence) while I code the other half with the maximum score of 2 (i.e. judicial independence). I multiplied the values of the terror threat variable by the judicial independence variable to code the interactive effect. Thereafter, I generated predicted probabilities of judicial confidence with standard errors on those predictions. Figure 1 presents the results of those predicted probabilities with 90% confidence intervals.

The results show a discernible interactive effect between judicial independence and terror threats. The effect of judicial independence on individual-level judicial confidence is discernible at the lowest levels of terror threats. Citizens in states with low or no terror threats are more likely to have confidence in their judicial system when it is independent compared to citizens in states with little to no terror threats and without an independent judiciary. This changes with increasing terror threats. Terror threats have the effect of raising citizen-level judicial confidence in the states without independent judiciaries as executive-led judicial policies (like military tribunals) have already co-opted judicial processes. Terror threats decrease citizen-level judicial confidence in states with independent judiciaries. 


Observe what happens as terror threats pass zero (i.e. the mean terror threat level in the EVS data) on the *x*-axis. Citizens in states without independent judiciaries have *more* confidence in their judicial systems than citizens in states with independent judiciaries and the confidence intervals do not overlap as terror threats increase. This is consistent with my hypothesis that citizens lose confidence in their judicial system and grow frustrated with the civil liberty protections they provide terror suspects under these conditions. Citizens interpret independent judiciaries capable of judicial review as providing unwanted legal protections for terror suspects and security threats.


# Conclusion, Implications


My manuscript starts with a puzzling observation about the lack of confidence British citizens profess in their judicial system. Multiple indicators suggest the British judicial system is among the most efficient and highest-quality judicial systems in the world and is free from subjugation to other state or societal actors. Yet, British confidence in the system lags behind judicial confidence elsewhere in Western Europe and British confidence in other state institutions. Why do British citizens express so little support for a judicial system as independent and civil liberty-oriented as the British model?

I argue that the independence of the British judiciary from other state and societal actors and its strength in protecting civil liberties may counterintuitively be part of the problem for British judicial confidence. Further, the United Kingdom has a long history of terror threats, unlike some of its Western European peers with higher levels of judicial confidence. The United Kingdom diverges from the "flexible", security-oriented judicial model that France has [@foley2013ctbf] and instead implements a rigorous system that provides legal protections of due process for terror suspects. As a result, citizens lose confidence in a judicial system that provides unwanted due process to suspected terrorists. My analysis extends the observation about the British puzzle to the rest of Europe using four waves of EVS data from 1982 to 2009 and finds a statistically significant negative interaction between terror threats and judicial independence on individual-level judicial confidence. 

This has important implications for how we understand the "liberty-security dilemma." This problem underlines many issues in current U.S. foreign policy <!--[@rosatiscott2014pusf] --> but it is a familiar problem for democratic governments and political science scholarship.<!--[^bay1958]--> My findings suggest multiple ways to think about this issue further. Most scholars and policymakers sympathetic to the restriction of judicial protections of liberty in the name of security couch their arguments on normative grounds. @posnervermeule2007tb argue that courts should defer to executives because courts are slow and rigid while the executive has an endowment advantage and the ability to act in secrecy for the cause of counter-terrorism. They even contend that errors that infringe on judicial process are natural and even desirable when compared with the alternatives proposed by staunch civil libertarians. @yoo2006wom, known for his infamous tenure as George W. Bush's Deputy Assistant U.S. Attorney General, authored numerous defenses of Bush-era policies ranging from the military prisons to the USA PATRIOT Act based on the nature of the combatant and the War on Terror. His defense was also normative in tone. More recent political science scholarship has countered this argument with formal models that show restrictions in liberty are not necessary for security [@dragu2011itt], that rigorous judicial review can lead to more informed policy-making even in the absence of policy expertise in the judiciary [@draguboard2015jrsp], and that judicial deference implied may lead to suboptimal counter-terror provisions [@bdm2007psp; @dragupolborn2014rlft].

My analysis illuminates the claims made by proponents of restricting civil liberties for national security, though my analysis is empirical and not normative. These executive-led policies that infringe on legal protections provided by an independent judiciary follow because citizens under terrorist duress look to the executive to provide counter-terrorism and provide for their security. My analysis suggests citizens seem to agree with @posnervermeule2007tb that courts are slow and rigid and should not be threatening judicial review when security is at stake. Public opinion may provide the leverage for state leaders to absorb some judicial processes into the executive under conditions of high terror threat.

<!-- [^bay1958]: @bay1958sf appears to be the first reference to his problem as we currently call it. -->

My findings further suggest the liberty-security dilemma is not unique to the United States. Democratic politicans and citizens wrestle with what constitutes an acceptable trade-off of individual liberty in pursuit of national security. The post-*Charlie Hebdo* reaction in France, for example, resembled the post-9/11 reactions in the United States and even Australia. The immediate fallout in France after *Charlie Hebdo* led to renewed energy on the "French Patriot Act," a controversial counter-terror bill patterned off the USA PATRIOT Act. Despite public opinion across Europe that was critical of the privacy infringements outlined the USA PATRIOT Act, citizens in Europe are prone to the same change in political attitudes when confronted with their own terror threats. This underlines the peculiar case of France. French citizens were among the most consistently critical citizens in the world for the terms of the USA PATRIOT Act as it applied to U.S. citizens [e.g. @prc2014gous]. After *Charlie Hebdo*, French citizens became open to a French version of the same controversial legislation [@jordan2015yougov].

One particularly troubling implication of my analysis concerns recent developments in the United States. On February 9, 2017, the United States Court of Appeals for the Ninth Circuit blocked key parts of a "travel ban" instituted by U.S. President Donald Trump via an executive order and refused a stay of the executive order upon appeal. Trump's reaction was troubling. He unleashed a series of tweets in response to these legal defeats that blasted the "so-called judge" and his "ridiculous" opinion for issuing a decision that "put our country in such peril" [@flores2017tt]. @trump2017syc also tweeted "the security of our nation is at stake" in all-caps in response to a follow-up legal defeat in which the same circuit court denied a stay of the executive order. Trump's behavior here lacked subtlety to experienced political commentators and political scientists [e.g. @blake2017tnss; @nyhan2017dal]. @posner2017wgsc captured this sentiment bluntly when he noted Trump's tweets are "clearly an attack on the independence of the judiciary." However, we should note these comments from Trump occur during a period of low terror threat in the United States and his behavior might be more cynical. @trump2017jcb himself states his intentions when he says "if something happens [i.e. a terror attack], blame him [the judge that first blocked the executive order] and [the] court system." The extent to which high terror threats decrease judicial confidence in countries with independent judiciaries, like the United States, creates a cynical incentive for state leaders to galvanize public opinion against an independent judiciary and pass blame for future terror attacks if the acquisition of more power is the goal.

My analysis has some final implications for the scholarly study of judicial independence and confidence in the judiciary. We know that these institutions are important stopgaps against democratic reversals. They also help facilitate peaceful transitions of power in new democracies and are a necessary institution that legitimates democratic experiments to citizens. They might even be a "necessary condition" to democracy [e.g. @howardcasey2003ijnd]. The implication is citizens have confidence in the judiciary when it is honest, diligent, and fair in its basic functions (i.e. when it is "independent"). My findings suggest this simple conclusion requires an important caveat. Citizens lose faith in their independent judiciary when security is at stake. By contrast, they gain faith in their judiciary when state leaders co-opt basic functions of judicial process in the name of counter-terrorism and security.

\newpage


# Acknowledgements

The author thanks---but cannot blame---Matthew Clary, Chris Kromphardt, Julia West, and the editors and three anonymous reviewers at *Political Research Quarterly* for their comments on previous versions of this manuscript. 

# Declaration of Conflicting Interests

The author declared no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.

# Funding

The author received no financial support for the research, authorship, and/or publication of this article.

# Supplemental Material

Replication files are available on the author's Github account ([github.com/svmiller](https://github.com/svmiller)).

